---
title: "Marriage Expectations"
author: "Joanna R. Pepin and Philip N. Cohen"
date: today
bibliography: references.bib
filters:
    - openlinksinnewpage
---

```{r notes}
#| echo: false
#| output: false
#| tbl-cap: ""

# Abstract: 200 words or less
# Main text: [approximately 500 words (excluding caption & references)]

```

# Abstract

Using [Monitoring the Future (MTF) Public-Use Cross-Sectional Datasets (1976-2022)](https://www.icpsr.umich.edu/web/ICPSR/series/35), this visualization presents U.S. twelfth graders marriage expectations. It shows declining optimism among young people in their belief that they will be a 'very good spouse' and that they will eventually get married. declines starting in XXXX, especially among young women.

### Keywords

marriage, attitudes, adolescents, gender, visualization

------------------------------------------------------------------------

\[motivation paragraph\]

We used data from the 1976 through 2022 waves of [Monitoring the Future (MTF) Public-Use Cross-Sectional Datasets](https://www.icpsr.umich.edu/web/ICPSR/series/35), an annual survey nationally representative of American students, [@miech2022MonitoringFutureContinuing] to examine young people's marital expectations. Figure 1 presents trends in U.S. twelfth graders' assessment of how good of a spouse they expect to be (panel A), whether they think they will get married (panel B), and the percentage point change in this belief over the last 10 years (panel C). Additional details are available in the [online supplement](https://jrpepin.github.io/MTF_Marriage/Supplementary%20Material.html) and data access and replication code are available at: <https://github.com/jrpepin/MTF_Marriage>.

Declining optimism in marital expectations is apparent in Figure 1. Only 38% of young men expected to be a 'very good' spouse in 1976 but by 2001, about 60% of young men expected their future selves would be a 'very good' spouse. Young women were more optimistic than young men, with 47% reporting their future selves' would be a 'very good' spouse in 1976, and their optimism peaked in 2012 (64%). Both young men and women's optimism has been declining since the early 2010s, with young women's expectations showing a slightly steeper decline than young men's expectations of being a 'very good' spouse. Still, about half of young men and women in 2022 believe their future selves will be a 'very good' spouse. Notably, as the proportion of youth who expect to be a 'very good' spouse declines, an increasing proportion of young men expect their future self will be 'poor,' whereas the proportion of young women who believe they will be a 'fairly good' spouse increased.

\[explain panel B findings here\]

\[explain panel C findings here\]

```{r setup}
#| echo: false
#| output: false
#| tbl-cap: ""
#| file: scripts/ME_00-1_setup and packages.R
```

```{r measures}
#| echo: false
#| output: false
#| tbl-cap: ""
#| file: scripts/ME_01_measures and sample.R
```

```{r figure-01}
#| echo: false
#| output: false
#| tbl-cap: ""
#| file: scripts/ME_02_figure.R
```

```{r figure-02, fig.width=9, fig.height=6.5}
#| echo: false
#| tbl-cap: ""

plbl

```

\[add summary findings & implications\]

# References
