---
title: "Marriage Expectations"
author: "Joanna R. Pepin and Philip N. Cohen"
date: today
bibliography: references.bib
filters:
    - openlinksinnewpage
---

```{r notes}
#| echo: false
#| output: false
#| tbl-cap: ""

# Abstract: 200 words or less
# Main text: [approximately 500 words (excluding caption & references)]

```

# Abstract

Using [Monitoring the Future (MTF) Public-Use Cross-Sectional Datasets (1976-2022)](https://www.icpsr.umich.edu/web/ICPSR/series/35), this visualization presents U.S. twelfth graders marriage expectations. It shows declining optimism among young people in their belief that they will be a 'very good spouse' and that they will eventually get married. declines starting in XXXX, especially among young women.

### Keywords

marriage, attitudes, adolescents, gender, visualization

------------------------------------------------------------------------

\[motivation paragraph\]

We used data from the 1976 through 2022 waves of [Monitoring the Future (MTF) Public-Use Cross-Sectional Datasets](https://www.icpsr.umich.edu/web/ICPSR/series/35), an annual survey nationally representative of American students, [@miech2022MonitoringFutureContinuing] to examine young people's marital expectations. Figure 1 presents trends in U.S. twelfth graders' assessment of how good of a spouse they expect to be (panel A), whether they think they will get married (panel B), and the percentage point change in this belief over the last 10 years (panel C). Additional details are available in the [online supplement](https://jrpepin.github.io/MTF_Marriage/Supplementary%20Material.html) and data access and replication code are available at: <https://github.com/jrpepin/MTF_Marriage>.

Figure 1 shows that after decades of growing optimism regarding marital expectations, attitudes have become more pessimistic in the last decade. About 38% of young men expected to be a 'very good' spouse in 1976, growing to about 60% by 2001 (panel A). Young women were more optimistic than young men, with 47% reporting their future selves' would be a 'very good' spouse in 1976, and their optimism peaked in 2012 (64%). Both young men and women's optimism has been waning since the early 2010s, with young women's expectations showing a slightly steeper decline than young men's expectations of being a 'very good' spouse. Notably, as the proportion of youth who expect to be a 'very good' spouse declines, an increasing proportion of young men expect their future self will be a 'poor' spouse, whereas the proportion of young women who believe they will be a 'fairly good' spouse has been growing. Still, in 2022, about half of young men and women believe their future selves will be a 'very good' spouse and about 80% of youth expect they will be a 'good' or 'very good' spouse.

A similar trend is apparent in young adults' expectations about whether they will eventually get married (panel B). In 1976, nearly 3 out of 4 young men and 4 out of 5 young women predicted their future selves would get married. At the peak, 84% of young men in 2007 and 88% of young women in 2005 expected to eventually get married. Beginning in the 2010s, a shrinking proportion of young people report they anticipate they will 'get married.' By 2022, about 3 out of 4 young men and women expected to eventually marry, a return to the share of young men in the mid-1970s and about 10% fewer young women than their 1976 counterparts. Instead, over the last decade, an increasing proportion of these young adults report they 'have no idea' if they will get married, and a minority of them increasingly report they expect they will never marry.

\[explain panel C findings here\]

```{r setup}
#| echo: false
#| output: false
#| tbl-cap: ""
#| file: scripts/ME_00-1_setup and packages.R
```

```{r measures}
#| echo: false
#| output: false
#| tbl-cap: ""
#| file: scripts/ME_01_measures and sample.R
```

```{r figure-01}
#| echo: false
#| output: false
#| tbl-cap: ""
#| file: scripts/ME_02_figure.R
```

```{r figure-02, fig.width=9, fig.height=6.5}
#| echo: false
#| tbl-cap: ""

plbl

```

\[add summary findings & implications\]

# References
