---
title: "Marriage Expectations"
author: "Joanna R. Pepin and Philip N. Cohen"
date: today
bibliography: references.bib
---

```{r notes}
#| echo: false
#| output: false
#| tbl-cap: ""

# Abstract: 200 words or less
# Main text: [approximately 500 words (excluding caption & references)]

```

# Abstract

Using [Monitoring the Future (MTF) Public-Use Cross-Sectional Datasets (1976-2022)](https://www.icpsr.umich.edu/web/ICPSR/series/35), this visualization presents U.S. twelfth graders marriage expectations. It shows declining optimism among young people in their belief that they will be a 'very good spouse' and that they will eventually get married. declines starting in XXXX, especially among young women.

### Keywords

marriage, attitudes, adolescents, gender, visualization

------------------------------------------------------------------------

\[motivation paragraph\]

We used data from the 1976 through 2022 waves of [Monitoring the Future (MTF) Public-Use Cross-Sectional Datasets](https://www.icpsr.umich.edu/web/ICPSR/series/35), an annual survey nationally representative of American students, [@miech2022MonitoringFutureContinuing] to examine young people's marital expectations. Figure 1 presents trends in U.S. twelfth graders' assessment of how good of a spouse they expect to be (panel A), whether they think they will get married (panel B), and the percentage point change in this belief over the last 10 years (panel C). Additional details are available in the [online supplement](https://jrpepin.github.io/MTF_Marriage/Supplementary%20Material.html) and data access and replication code are available at: <https://github.com/jrpepin/MTF_Marriage>.

\[findings\]

```{r setup}
#| echo: false
#| output: false
#| tbl-cap: ""
#| file: scripts/ME_00-1_setup and packages.R
```

```{r measures}
#| echo: false
#| output: false
#| tbl-cap: ""
#| file: scripts/ME_01_measures and sample.R
```

```{r figure-01}
#| echo: false
#| output: false
#| tbl-cap: ""
#| file: scripts/ME_02_figure.R
```

```{r figure-02, fig.width=9, fig.height=6.5}
#| echo: false
#| tbl-cap: ""

plbl

```

\[more findings\]

# References
