---
title: "Marriage Expectations"
author: "Joanna R. Pepin and Philip N. Cohen"
date: today
---

```{r notes}
#| echo: false
#| output: false
#| tbl-cap: ""

# Abstract: 200 words or less
# Main text: [approximately 500 words (excluding caption & references)]

```


# Abstract

Using [Monitoring the Future (MTF) Public-Use Cross-Sectional Datasets (1976-2022)](https://www.icpsr.umich.edu/web/ICPSR/series/35), this visualization presents U.S. twelfth graders marriage expectations by gender, racial group, and mothers' education. It shows...


### Keywords

marriage, attitudes, adolescents, visualization

---

[motivation paragraph]

[data/methods]
Additional details are available in the [online supplement](https://jrpepin.github.io/MTF_Marriage/Supplementary Material.html) and data access and replication code are available at: [https://github.com/jrpepin/MTF_Marriage](https://github.com/jrpepin/MTF_Marriage).

[findings]


```{r setup}
#| echo: false
#| output: false
#| tbl-cap: ""
#| file: scripts/ME_00-1_setup and packages.R
```

```{r measures}
#| echo: false
#| output: false
#| tbl-cap: ""
#| file: scripts/ME_01_measures and sample.R
```

```{r figure-01}
#| echo: false
#| output: false
#| tbl-cap: ""
#| file: scripts/ME_02_figure.R
```

```{r figure-02, fig.width=6.5, fig.height=8}
#| echo: false
#| tbl-cap: ""

ptext

```

[more findings]

# References 
