---
title: "Marriage Expectations"
author: "Joanna R. Pepin and Philip N. Cohen"
date: today
bibliography: references.bib
filters:
    - openlinksinnewpage
---

```{r notes}
#| echo: false
#| output: false
#| tbl-cap: ""

# Abstract: 200 words or less
# Main text: [approximately 500 words (excluding caption & references)]

```

# Abstract

Using [Monitoring the Future (MTF) Public-Use Cross-Sectional Datasets (1976-2022)](https://www.icpsr.umich.edu/web/ICPSR/series/35), this visualization presents U.S. twelfth graders marriage expectations. It shows declining optimism among young people in their belief that they will be a 'very good spouse' and that they will eventually get married. declines starting in XXXX, especially among young women.

### Keywords

marriage, attitudes, adolescents, gender, visualization

------------------------------------------------------------------------

\[motivation paragraph\]

<https://www.bgsu.edu/ncfmr/resources/data/family-profiles/brown-high-school-seniors-expectation-to-marry-2020-fp-22-04.html>

<https://link.springer.com/article/10.1007/s11113-023-09833-8>

We used data from the 1976 through 2022 waves of [Monitoring the Future (MTF) Public-Use Cross-Sectional Datasets](https://www.icpsr.umich.edu/web/ICPSR/series/35), an annual survey nationally representative of American students, [@miech2022MonitoringFutureContinuing] to examine young people's marital expectations. Figure 1 presents trends in U.S. twelfth graders' assessment of how good of a spouse they expect to be (panel A), whether they think they will get married (panel B), and the percentage point change in this belief over the last 10 years (panel C). Additional details are available in the [online supplement](https://jrpepin.github.io/MTF_Marriage/Supplementary%20Material.html) and data access and replication code are available at: <https://github.com/jrpepin/MTF_Marriage>.

After decades of growing optimism regarding marital expectations, attitudes have become more pessimistic in the last decade (Figure 1). About 38% of young men expected to be a 'very good' spouse in 1976, growing to about 60% by 2001 (panel A). Young women were more optimistic than young men, with 47% reporting their future selves' would be a 'very good' spouse in 1976, and their optimism peaked in 2012 (64%). Optimism has been waning since the early 2010s, with young women's expectations showing a steeper decline than young men's expectations of being a 'very good' spouse. An increasing proportion of young men expect their future self will be a 'poor' spouse, whereas the proportion of young women who believe they will be a 'fairly good' spouse has been growing. Still, in 2022, about 80% of youth expect they will be a 'good' or 'very good' spouse.

A similar trend is apparent regarding expectations about whether they will eventually get married (panel B). In 1976, nearly 3 out of 4 young men and 4 out of 5 young women predicted their future selves would 'get married.' Thereafter, a shrinking proportion of young people reported they anticipated getting married. By 2022, about 3 out of 4 young men and women expected to eventually marry, a return to the share of young men in the mid-1970s and about 10% fewer young women than their 1976 counterparts. Panel C shows the decline in expectations of getting married has been twice as large for young women (-14%) than for young men (7%). In the last decade, young people were increasingly likely to report they were unsure if they will marry, whereas expecting not to marry remained rare.

```{r setup}
#| echo: false
#| output: false
#| tbl-cap: ""
#| file: scripts/ME_00-1_setup and packages.R
```

```{r measures}
#| echo: false
#| output: false
#| tbl-cap: ""
#| file: scripts/ME_01_measures and sample.R
```

```{r figure-01}
#| echo: false
#| output: false
#| tbl-cap: ""
#| file: scripts/ME_02_figure.R
```

```{r figure-02, fig.width=9, fig.height=6.5}
#| echo: false
#| tbl-cap: ""

plbl

```

\[add summary findings & implications\]

<https://x.com/JonHaidt/status/1734931511168712808>

# References
